---
title: "Data preparation and analysis"
output:
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)

# Packages used
library(readr)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(e1071)
library(tidytext)
library(DT)
library(tm)
library(tidyverse)
library(stringr)
library(leaflet)
library(ggmap)
library(knitr)
library(Hmisc)
library(scales)

reviews<-read_csv("reviews.csv")
calendar<-read_csv("calendar.csv")
listings<-read_csv("listings.csv")
```

### Listing price

```{r}
#calendar has unformatted price data
calendar$price<-as.numeric(gsub('[$,]','',calendar$price))
summary(calendar$price)
```

```{r}

#listings has unformatted data as well
listings$price<-as.numeric(gsub('[$,]','',listings$price))
summary(listings$price)
```
```{r}
#remove na values
c<-which((is.na(calendar$price)))
calendar<-calendar[-c,] #1305030
calendar$day<-weekdays(calendar$date)
calendar$month<-month(calendar$date)
stats<- calendar %>% summarise(mean=mean(price), median=median(price), stdDev=sd(price), q1=quantile(price,probs=0.25), q3=quantile(price, probs=0.75),n=n())
```

mean
208.501
median
125
stdDev 
340.2615
q1 
80
q3
215
n
1305030

### Outlier detection for listing price

```{r}
#outlier detection for price
skewness(calendar$price) #11
iqr<-stats$q3 - stats$q1
iqrm<-1.5*iqr
od<-stats$q3 + iqrm
ecdf(calendar$price)(od)
#90% of the data is retained after outlier detection

cleaned_calendar<-calendar %>%
  filter(price<od)
od
skewness(cleaned_calendar$price) #1.1

by_month<-calendar %>%
  group_by(month) %>%
  summarise(avg_price=mean(price))

by_day<-calendar %>%
  group_by(day) %>%
  summarise(avg_price=mean(price))

```

The technique of outlier detection employed here is the 1.5IQR rule, after which approx. 90% of the data is retained which is a pretty satisfactory representation.

After cleaning the calendar for outliers, the price can be looked at from a monthly and day granularity level.
The month of June saw the highest average price of around $220 while February saw the least of around $189


```{r}
#Finding min and max values in the calendar data
min(calendar$date)
max(calendar$date)
```
Another point to note is that the price and availability data is available for approx. a year i.e. from November 15 2018 to November 20 2019


### Most Frequent words used by hosts in descriptions
```{r}
# unnest_tokens function to tokenise
listings_words <- listings %>%
  select(id, description, price, review_scores_accuracy, review_scores_rating) %>%
  unnest_tokens(word, description) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "^[a-z']+$"))

#plot the graph
common_listings <- listings_words %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  top_n(n = 20, wt = count) %>%
  ggplot() +
  geom_bar(mapping = aes(x=reorder(word, count),
                         y=count),
           stat="identity", fill = "blue") +
  labs(title="Top 20 words described in listings",
       y="Word count", x="Most common Words") +
  coord_flip() +
  theme_minimal()
  
common_listings

```

From the plot above of top 20 words in listing descriptions, it seems like most of the hosts mention about the proximity to metro in the description.

### Most Frequent words used by guests in reviews
```{r}
# Using unnest_tokens function to takenise
review_words <- reviews %>%
  unnest_tokens(word, comments) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "^[a-z']+$"))

op <- par(mar = c(9,4,4,2) + 0.1)
#plot the graph
common_reviews <- review_words %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  top_n(n = 20, wt = count) %>%
  ggplot() +
  geom_bar(mapping = aes(x=reorder(word, count), y=count),
           stat="identity", fill = "orange") +
  coord_flip() +
  labs(title="Top 20 words described in Reviews",
       y="Word count", x="Words") +
  theme_minimal()+ scale_y_sqrt(labels = scales::comma)
common_reviews
par(op)
```

From the above plot of top 20 keywords in reviews, it seems like stay, location, clean, host, comfort are some of the important factors that matter to the guests.

### Neighbourhood

```{r}
factpal <- colorFactor(topo.colors(3), listings$neighbourhood_cleansed)

popup <- paste0("<strong>'hood: </strong>", listings$neighbourhood_cleansed)

leaflet(listings) %>% addProviderTiles("CartoDB.DarkMatter") %>%
  addCircleMarkers(
    color = ~factpal(neighbourhood_cleansed),
    stroke = FALSE, fillOpacity = 0.5, radius = 1.2,
    popup = ~popup
  )
```




```{r}

#Finding the count by group_by and sort
listing_groupby<- listings %>%
  group_by(neighbourhood_cleansed)
count_list <- count(listing_groupby, sort = TRUE)
kable(count_list)
```

From the plot and table above, we see that most number of listings are close to the neighbourhoods Columbia Heights, Union stations, Capitol Hill etc.


### Review score rating
```{r}

review_desc <- listings$review_scores_rating
summary(review_desc )
```
As seen above, most of the guest who review give high scores.

### Different listings based on Room type 

```{r}
room_groupby<- listings %>%
  group_by(room_type)
count_room <- count(room_groupby, sort = TRUE)
kable(count_room)

```

### Property Type of listings

```{r}
listings$property_type = ifelse(listings$property_type == "Apartment", "Apartment", 
                         ifelse(listings$property_type == "Bed & Breakfast","B&B",
                         ifelse(listings$property_type == "Condominium","Condominium",
                         ifelse(listings$property_type == "House","House",
                         ifelse(listings$property_type == "Loft","Loft",
                         ifelse(listings$property_type == "Townhouse","Townhouse",
                         ifelse(listings$property_type == "Dorm","Dorm", "Other")))))))
listings$property_type = as.factor(listings$property_type)
```

We have only kept Apartment, b&b, Condominium, House, Loft, Townhouse, and Dorm in Property.Type and the rest would be categorised to Others.

```{r}
s <- unique(listings$property_type)

cr <- data.frame(count_room)
piepercent<- round(100*(cr$n/sum(cr$n)),2)
bp<- ggplot(cr, aes(x="", y=n, fill=room_type))+
geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y")+geom_text(label=piepercent)+theme(axis.text.x=element_blank())
pie
op <- par(mar = c(9,4,4,2) + 0.1)
barplot(table(listings$property_type),col = "brown",las=2)
par(op)
```

The above pie chart shows people prefer entire property than private rooms or shared rooms. On exploring further about property type, it is seen that majority of the listings are Apartment, House, townhouse.




