---
title: "Airbnb"
output: word_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)

# Packages used
library(readr)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(e1071)
library(tidytext)
library(DT)
library(tm)
library(tidyverse)
library(stringr)
library(leaflet)
library(ggmap)
library(knitr)
library(Hmisc)
library(scales)

reviews<-read_csv("reviews.csv")
calendar<-read_csv("calendar.csv")
listings<-read_csv("listings.csv")
```

## Exploratory analysis


### Listing price

```{r}
```

### Calender
```{r}
#Finding min and max values in the calendar data
min(calendar$date)
max(calendar$date)
```
The calendar covers 365 days, that is, price and availability every day for one year. In our case, from 2018/11/15 to 2019/11/20.


### Most Frequent words used by hosts in descriptions
```{r}
# unnest_tokens function to obtain one-row-per-term-per-listing-description
listings_words <- listings %>%
  select(id, description, price, review_scores_accuracy, review_scores_rating) %>%
  unnest_tokens(word, description) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "^[a-z']+$"))

#plot the graph
common_listings <- listings_words %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  top_n(n = 20, wt = count) %>%
  ggplot() +
  geom_bar(mapping = aes(x=reorder(word, count),
                         y=count),
           stat="identity", fill = "blue") +
  labs(title="Top 20 words described in listings",
       y="Word count", x="Most common Words") +
  coord_flip() +
  theme_minimal()
  
common_listings
```

From the plot above of top 20 words in listing descriptions, it seems like most of the hosts mention about the proximity to metro in the description.

### Most Frequent words used by guests in reviews
```{r}
#We need to use the unnest_tokens function to obtain one-row-per-term-per-listing-description
review_words <- reviews %>%
  unnest_tokens(word, comments) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "^[a-z']+$"))

#plot the graph
common_reviews <- review_words %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  top_n(n = 20, wt = count) %>%
  ggplot() +
  geom_bar(mapping = aes(x=reorder(word, count), y=count),
           stat="identity", fill = "orange") +
  coord_flip() +
  labs(title="Top 20 words described in Reviews",
       y="Word count", x="Words") +
  theme_minimal()

common_reviews
```

From the above plot of top 20 keywords in reviews, it seems like stay, location, clean, host, comfort are some of the important factors that matter to the guests.


```{r}
factpal <- colorFactor(topo.colors(3), listings$neighbourhood_cleansed)

popup <- paste0("<strong>'hood: </strong>", listings$neighbourhood_cleansed)

leaflet(listings) %>% addProviderTiles("CartoDB.DarkMatter") %>%
  addCircleMarkers(
    color = ~factpal(neighbourhood_cleansed),
    stroke = FALSE, fillOpacity = 0.5, radius = 1.2,
    popup = ~popup
  )
```




```{r}

#Finding the count by group_by and sort
listing_groupby<- listings %>%
  group_by(neighbourhood_cleansed)
count_list <- count(listing_groupby, sort = TRUE)
kable(count_list)
```

From the plot and table above, we see that most number of listings are close to the neighbourhoods Columbia Heights, Union stations, Capitol Hill etc.


### Review score rating
```{r}

review_desc <- listings$review_scores_rating
summary(review_desc )
```
As seen above, most of the guest who review give high scores.

### Different listings based on Room type 

```{r}
room_groupby<- listings %>%
  group_by(room_type)
count_room <- count(room_groupby, sort = TRUE)
kable(count_room)

```

```{r}

s <- unique(listings$property_type)

cr <- data.frame(count_room)
piepercent<- round(100*(cr$n/sum(cr$n)),2)
bp<- ggplot(cr, aes(x="", y=n, fill=room_type))+
geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y")+geom_text(label=piepercent)+theme(axis.text.x=element_blank())
pie

barplot(table(listings$property_type),col = "brown",las=2)
```

The above pie chart shows people prefer entire property than private rooms or shared rooms. On exploring further about property type, it is seen that majority of the listings are Apartment, House, townhouse.
